---
title: "Example Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

Doctor Who is an extremely long-running British television program. The show was revived in 2005, and has proven very popular since then. "Doctor Who" is a British science fiction television series that first premiered in 1963. The show follows the adventures of the Doctor, a Time Lord from the planet Gallifrey who travels through time and space in the TARDIS, a time machine that looks like a British police box. The Doctor, often accompanied by companions, explores different worlds, encounters a variety of aliens, and seeks to right wrongs and help those in need. One of the show's defining features is the Doctor's ability to regenerate, allowing for different actors to take on the role. "Doctor Who" has become a cultural phenomenon, with a rich history spanning classic and modern eras, and it continues to captivate audiences with its imaginative storytelling and iconic characters.To celebrate this year's 60th anniversary of Doctor Who, the authors of this dataset presented three datasets.

The data this week comes from Wikipedia's [List of Doctor Who episodes]([https://en.wikipedia.org/wiki/List_of_Doctor_Who_episodes\_(2005%E2%80%93present)](https://en.wikipedia.org/wiki/List_of_Doctor_Who_episodes_(2005%E2%80%93present)){.uri} via the {datardis} package by Jonathan Kitt.

In this analysis we try to understand the effect of 'specials' on the ratings of a show(as doctor who has many special episodes). We also analyse the changes in durations of the show over time, its effect on the ratings. We further compartmentalize the results to get a better idea of the underlying phenomenon.

```{r setup calculate_CI}
library(transport)
library(tidytuesdayR)
library(ggplot2)
library(readr)
library(purrr)

#install.packages("tidytuesdayR")

 #tuesdata <- tidytuesdayR::tt_load('2023-11-28')
 # ## OR
 #tuesdata <- tidytuesdayR::tt_load(2023, week = 39)

#  drwho_episodes <- tuesdata$drwho_episodes
# drwho_directors <- tuesdata$drwho_directors
# drwho_writers <- tuesdata$drwho_writers

     # dir.create("data")
     #
     # saveRDS(richmondway, file = "data/richmondway.RDS")

    #my_data <- readRDS("~Anaghpackage/data/richmondway.RDS")

 drwho_episodes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-11-28/drwho_episodes.csv')
 drwho_directors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-11-28/drwho_directors.csv')
 drwho_writers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-11-28/drwho_writers.csv')

#
    my_data1= drwho_episodes
    my_data2= drwho_directors
    my_data3= drwho_writers
```

We save the data below-

```{r setup, include=FALSE}
# Load necessary libraries and set options
library(fs)  # For file system operations

# Function to check and create directory if not exists
create_data_directory <- function() {
  data_dir <- "data"
  
  if (!dir_exists(data_dir)) {
    dir_create(data_dir)
    cat("Data directory created.\n")
  } else {
    cat("Data directory already exists.\n")
  }
}

# Function to save data
save_data <- function(data1) {
  data_len=deparse(substitute(data1))
  data_file=paste("data/",data_len,".csv",sep="")
  write.csv(data1, data_file, row.names = FALSE)
  cat("Data saved to", data_file, "\n")
}

# Function to read data
read_data <- function(data1) {
  data_len=deparse(substitute(data1))
  data_file=paste("data/",data_len,".csv",sep="")
  
  if (file_exists(data_file)) {
    data <- read.csv(data_file)
    cat("Data read from", data_file, "\n")
    return(data)
  } else {
    cat("Data file not found. Please save the data first.\n")
    return(NULL)
  }
}


```

The data directory is created in Vignettes.
```{r 1}
create_data_directory()

if (!file.exists("data/my_data3.csv")) {
  save_data(my_data3)
} else {
  my_data3 <- read_data(my_data3)
}

if (!file.exists("data/my_data2.csv")) {
  save_data(my_data2)
} else {
  my_data2 <- read_data(my_data2)
}

if (!file.exists("data/my_data1.csv")) {
  save_data(my_data1)
} else {
  my_data1 <- read_data(my_data1)
}
```

```{r Data Visualization}

#read the dataset
library(dplyr)
library(lubridate, warn.conflicts = FALSE)
library(forcats)
library(tidyr)

# We summarise the data and get a glimpse of the data set using the two functions below respectively.
summarise(my_data1)
glimpse(my_data1)

#This functions counts the numbers of unique value corresponding to the dataset, and it helps us understand the presence of any repetitions.
count(my_data1)

#We use the distinct function to find distinct value in the data, and convert the data structure to tibble for better operationality.
my_data_new1=distinct(my_data1)
my_data_new1=as_tibble(my_data_new1)
my_data_new1=arrange(my_data_new1)

# We summarise the data and get a glimpse of the data set using the two functions below respectively.
summarise(my_data2)
glimpse(my_data2)

#This functions counts the numbers of unique value corresponding to the dataset, and it helps us understand the presence of any repetitions.
count(my_data2)

#We use the distinct function to find distinct value in the data, and convert the data structure to tibble for better operationality.
my_data_new2=distinct(my_data2)
my_data_new2=as_tibble(my_data_new2)
my_data_new2=arrange(my_data_new2)

# We summarise the data and get a glimpse of the data set using the two functions below respectively.
summarise(my_data3)
glimpse(my_data3)

#This functions counts the numbers of unique value corresponding to the dataset, and it helps us understand the presence of any repetitions.
count(my_data3)

#We use the distinct function to find distinct value in the data, and convert the data structure to tibble for better operationality.
my_data_new3=distinct(my_data3)
my_data_new3=as_tibble(my_data_new3)
my_data_new3=arrange(my_data_new3)

#We also find the current time to be able to understand the date till which the analysis is valid. The analysis might not be valid at a future date.
Sys.timezone()

now(tzone = "America/New_York")
```

We now use Purr to simultaneously apply functions to vectors.

```{r purrr}
library(purrr)
vec_req <- as.numeric(my_data_new1$rating)
result <- map_dbl(vec_req, ~(. - min(vec_req)) / (max(vec_req) - min(vec_req)))
#We scale the ratings to 0,1
result

#We convert all the uppercase names of the directors to lower case.
lowercase_words <- map_chr(my_data_new2$director, tolower)
lowercase_words
```

The scaling gives us novel insights about the distribution of the ratings and we see that the ratings are fairly homogeneous.

We now apply the Optimal Transport maps to find the distance between the ratings in regular and special episodes.

```{r optimal}
library(dplyr)
library(transport)
datarate1=as.matrix((my_data_new1%>%filter(my_data_new1$type==unique(my_data_new1$type)[1]))$rating)
datarate2=as.matrix((my_data_new1%>%filter(my_data_new1$type==unique(my_data_new1$type)[2]))$rating)

num_1=length(datarate1)
num_2=length(datarate2)
numc=num_1*num_2
nums=num_1+num_2

ground_p <- 2
p <- 1
w1 <- rep(1/num_1, num_1)
w2 <- rep(1/num_2, num_2)

cost_matrix_Lp<-function(x,y,ground_p){
  mat=matrix(0,nrow(x),nrow(y))
  for(i in 1:nrow(x)){
    for(j in 1:nrow(y)){
      mat[i,j]<-(sum((x[i,] - y[j,])^ground_p))^(1/ground_p)
    }
  }
  return(mat)
}
costm1=cost_matrix_Lp(datarate1, datarate2, ground_p=2)
#The transport map is found below
res <- transport(w1, w2, costm=costm1)
res

#The wasserstein distance is found below
wasserstein(a=w1, b=w2, costm=cost_matrix_Lp(datarate1,datarate2,ground_p=2))
```
The transport map and low wasserstein distance shows that the two rating distributions are not far from each other.

We firstly look at the progression of seasons/episodes and correlate it with the duration of the episodes. We can see that there has been an increase as the seasons progressed. However most of the episodes are 45 minutes long.

```{r plot 1}
ggplot(my_data_new1, aes(x = production_code, y = duration, group=1)) +
  geom_point() +  geom_line() +
  labs(
    title = "Season-Episode vs Duration",
    subtitle = "For Doctor Who",
    caption = "Data source: Tidytuesday"
  )

```

We also looked at the relationship between the progression of seasons with the ratings of the episode. There is a tendency of episodes to have a similar ratings, but there are periods irrespective of type of episodes, where the ratings have fluctuated a lot.

```{r plot 2}

ggplot(my_data_new1, aes(x = production_code, y = rating, group=1)) +
  geom_point() +  geom_line() + facet_wrap(~ type)+
  labs(
    title = "Season-Episode vs Rating",
    subtitle = "For Doctor Who",
    caption = "Data source: Tidytuesday"
  )


```

Finally we look at the effect of type of season on the ratings. It seems that there is not much effect of type of episodes.

```{r plot 3}
ggplot(my_data_new1, aes(x = type, y = rating)) +
  geom_boxplot() +
  labs(
    title = "Episode Type vs Rating",
    subtitle = "For Doctor Who",
    caption = "Data source: Tidytuesday"
  )

```

In this analysis, we sought to comprehensively examine the impact of 'specials' on the ratings of a show, taking into consideration the numerous special episodes featured, particularly evident in series such as Doctor Who. Additionally, we delved into the variations in the show's duration over time, exploring how these changes influenced audience ratings. To enhance our understanding of the underlying phenomenon, we further segmented the results, providing a nuanced and detailed perspective on the observed patterns.

Packages used- summarise, glimpse, count, distinct, as_tibble, arrange, Sys.timezone now

Purrr- map_dbl, map_chr

Transport- transport, Wasserstein

Geom- point, line, boxplot

facet_wrap
